---
title: "Prediction Assignment"
author: "Paul Kavitz"
date: "April 3, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(caret)
library(randomForest)
library(MASS)
library(rpart)
setwd("Z:/Kliluk/Coursera/C8W4")
```

### Executive Summary
This analysis predicts whether a participant correctly or incorrectly performed a barbell lift using data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.

### Injest, Clean, and Create Data Sets

```{r echo=TRUE, message=TRUE}
set.seed(1001)
base <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!",""))
testing <- read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!",""))
# Exploratory Analysis
# dim(base); dim(testing); summary(base); str(base); head(base)

# Remove attribute columns that are not legitimate predictors
base <- base[,-c(1:7)]
testing <- testing[,-c(1:7)]

# Delete columns where all data is NA
base <- base[,colSums(is.na(base))==0]
testing <- testing[,colSums(is.na(testing))==0]

inTrain <- createDataPartition(base$classe, p = 3/4, list=FALSE)
training <- base[inTrain,]  ## training data set
validation <- base[-inTrain,] ## Cross-validation data set
```

### Prediction Model 1: Random Forest
Random Forest method is applied for the first prediction model, trained in the training set and evaluated in the validation set.
```{r echo=TRUE}
mod_rf <- randomForest(classe ~ ., data=training, method="class")
pred_rf <- predict(mod_rf, validation, type="class")
confusionMatrix(pred_rf, validation$classe)
```
From the confusion matrix we can see that the accuracy of the Random Forest model in our validation test set is 0.9949.

### Prediction Model 2: Linear Discriminant Analysis
Linear Discriminant Analysis method is applied for the second prediction model, trained in the training set and evaluated in the validation set.
```{r echo=TRUE}
mod_lda <- train(classe ~ ., data=training, method="lda")
pred_lda <- predict(mod_lda, validation)
confusionMatrix(pred_lda, validation$classe)
```
From the confusion matrix we can see that the accuracy of the LDA model in our validation test set is 0.7153.

### Prediction Model 3: Classification Tree
Classification Tree method is applied for the third prediction model, trained in the training set and evaluated in the validation set.
```{r echo=TRUE}
mod_class <- rpart(classe ~ ., data=training, method="class")
pred_class <- predict(mod_class, validation, type="class")
confusionMatrix(pred_class, validation$classe)
```
From the confusion matrix we can see that the accuracy of the LDA model in our validation test set is 0.7508.

### Conclusion
The Random Forest method yields the best accuracy of the three assessed models.  This model is now applied to the testing set per assignment requirements to assess model quality.
```{r echo=TRUE}
pred_rft <- predict(mod_rf, testing, type="class")
n=length(pred_rft)
for(i in 1:n){
    write.table(pred_rft[i],file=paste0("p",i,".txt"),
                quote=FALSE, row.names=FALSE, col.names=FALSE)
}
```